\textbf{Pervasive Edge: Functions Everywhere}

Full transparency is not possible

\textbf{Private Edge vs Public Edge}

Different access and usage policies

\textbf{Indoor Edge vs Mobile Edge}

Different discovery protocol

\textbf{Critical Edge vs Opportunistic Edge}

Difference acquisition and allocation policies


\textbf{On Demand vs On Premise}

\textbf{Internet Edge vs LAN Edge}


~~~

Open Questions

Should A3-E encompass other execution models than FaaS? If yes, how do we


the main problem now is the following: A3-E now covers the whole continuum. From the domain side,A3-E process was designed to tackle the problem of efficiency of edge domains. The description of the process is still reflecting this, including the use of the four phases and FaaS, whereas cloud and mobile may use a simplified. To address this, I will break the proposal into parts reflecting the continuum: cloud, edge, and mobile. Each part should have its own description of the A3-E process. 


guys, does it make sense for an edge domain to be "unavailable" for an application because of lack of resources? I'm adding examples to the proposal and one idea was to have critical applications with high availability and priority over the domain resources. But for me it looks odd to say that non-critical applications would be able to use a given edge domain only if "resources are available". For instance, if you pay for 4G, you would be angry if 

~~~

federated clouds are not addressed, nor the more complex cases of intra and inter-domain placement