\subsection{Domain Manager: Local-Edge}\label{sec:local-edge-domain-DSM}

%\subsubsection{Local-Edge Domains}

The domain manager prototype\footnote{Documentation and source code available at \url{https://github.com/deib-polimi/A3-E-DSM-local-edge/}} described herein focuses on a local-edge domain, and encompasses A3-E's Awareness and Acquisition. Both the Allocation and Engagement activities are delegated to the FaaS platform (OpenWhisk), which allocates $\mu$-services to its pool of containers and handles client application requests (fired by the middleware) by means of RESTful endpoints.

The local-edge manager prototype implements Awareness by broadcasting \textit{domain identification} UDP signals in a constant interval. A client device entering the network replies --- by means of a UDP unicast --- with the \textit{client identification} signal containing the $\mu$-services its client application requires, along with the respective repository from which $\mu$-service artifacts can be fetched during Acquisition (as described in Sec.~\ref{sec:A3-E-awareness}). 

For each identified $\mu$-service, the manager proceeds with Acquisition. Among the downloaded files, a descriptor provides instructions regarding installation (e.g., compilation of Java classes and required dependencies). In particular, the manager prototype relies on Gradle\footnote{https://gradle.org/}, a state-of-art build tool commonly employed with projects ranging from mobile applications to $\mu$-services.

Once downloaded and built, the edge domain Acquisition finishes with the deployment of $\mu$-service function(s) and dependences to OpenWhisk by means of its command line interface and the generation of a \textit{$\mu$-service acquired} signal with the same UDP unicast channel. In case of failure, the mobile middleware is informed with a \textit{$\mu$-service denied} signal (as described in Sec.~\ref{sec:A3-E-acquisition}).

\subsection{Domain Manager: Mobile Domain}\label{sec:mobile-domain-DSM}

Herein we describe a domain manager implementation for the Android platform\footnote{Documentation and source code available at \url{https://github.com/deib-polimi/A3-E-DSM-mobile/}}. 
%As previously explained in Sec.~\ref{sec:A3-E}, mobile domains are exempt of performing Acquisition, thus this section details the implementation of A3-E's Awareness and Engagement. 
%[Danilo] shall we divise the middleware from the mobile domain or are they packed together?
The resulting implementation was packaged as a module within the mobile middleware for Android platform (described in Sec.~\ref{sec:mobile_middleware}). 

The prototype implements Awareness by triggering a system-level \textit{domain identification} signal once it has been loaded by the middleware and by listening to a \textit{client identification} reply signal. 
In contrast with cloud and edge domains, 
%Instead of a repository, 
each \textit{$\mu$-service identified} signal contains the qualified name (e.g., the system path of a Java class implementing the static function), which is added to a \textit{service registry} implementing Acquisition. 

The prototype supports two types of $\mu$-service functions: Java functions, which are natively supported, and JavaScript functions, which require a JNI wrapper for their execution by the Android platform. Note that existing FaaS platforms support a variety of other languages and runtimes. More comprehensive implementations of the mobile domain can either make use of additional wrappers; developers may also provide native implementations of $\mu$-service functions for this domain. 

%for each $\mu$-service required by the application. 

Once a $\mu$-service function is registered, the mobile domain triggers a \textit{$\mu$-service acquired} event, which enables its selection by the middleware. During Engagement and upon the selection of this domain, \textit{$\mu$-service request} events are handled with the lookup of the corresponding function. Once found, the function is called with the parameters composing the original \textit{C-request}; when finished, a \textit{$\mu$-service response} signal containing the execution result is generated.

 %Decoupling between the client application and its mobile domain was achieved with the use of Android events triggered by the CSM (see Section~\ref{sec:CSM}) and handled by the DSM. 
%Each request contains the unique name of a microservice. 


